---
title: "Cleaning Data"
author: "Team 7"
date: "2022-09-02"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```


```{r}
#Libraries used
library(tidyverse)
library(janitor)
library(dplyr)
```

Reading the data from .csv file
```{r}
df <- read_csv("D:/MS_in_DS/Stats_1/Assignments/Assignment_1/Nikhil_BM_Stats_Assignment1/Stats_Assignment_1_Dirty.csv")
```
Displaying the data
```{r}
df
```
We noticed that the column names are not in correct format.
Renaming the columns using function clean_name from janitor package

```{r}
df <- 
df %>%
  clean_names()
```

```{r}
colnames(df)
```
In this data song_id will work as Primary key if required so sr_no is not required

Removing unnecessary columns from the data 
```{r}
df = subset(df, select = -sr_no )
col_names <- colnames(df)
```

Checking for NULL values in all columns
```{r}
df %>% summarise(across(everything(), ~ sum(is.na(.))))
```
From the above table we can infer that columns like song_name,
track_number, explicit, danceability are not dependent on any other column also replacing danceability with it's mean might not give desired results hence it is safer to remove them

But artist_name might be dependent on artist_id
Also we can substitute disc_number with the most frequent disc number

Substituting artist name on basis of artist_id
Creating a new table just for artist_id and artist name

```{r}
df1 <- df[c("artist","artist_id")]
#Removing redundancy
df1 <- unique(df1)
#Deleting all rows that contain NA
df1 <- na.omit(df1)
df1
```
Joining the 2 data frames based on artist_id
```{r}
df <- 
df %>%
  left_join(df1,by = "artist_id")

#Removing artist_x
df <-  subset(df, select = -artist.x )
#Renaming artist.y to artist
df <- rename(df, artist = artist.y)

#Changing the sequence of columns
df <- df %>% select(song_id:album_id,'artist',everything())
```
```{r}
df
```

We also see that,Disc Number column has NA. 
So replacing NA with Mode in case of disc_number.
```{r}
df <- 
df %>% 
  mutate(disc_number = ifelse(is.na(disc_number),
                            mode(disc_number),
                            disc_number))
```

We also see song names has null values. These rows where we don't have
song name details cannot be replaced by any other qualitative data. 
Hence removing all rows that contain NA

```{r}
df <- na.omit(df)
distinct(df)
```
```{r}
df %>% summarise(across(everything(), ~ sum(is.na(.))))
```
Finally we have no NA values in our dataframe



Having a look at column Sub genre and Main Genre
```{r}
unique(df$genre)
```
We notice that there are some unnecessary character's in genre name
like [,',]
So removing them
```{r}
# to remove "['" in genre column at the front
df$sub_genre <- str_remove(df$sub_genre, pattern = "^\\[\\'")
# to remove "]" in genre column at the back
df$sub_genre  <- gsub("]","",as.character(df$sub_genre))

# to remove "['" in genre column at the front
df$main_genre <- str_remove(df$main_genre, pattern = "^\\[\\'")
# to remove "]" in genre column at the back
df$main_genre  <- gsub("]","",as.character(df$main_genre))
```
Once we have removed the extra characters, the sub_genre and main_genre column is clean. 
```{r}
unique(df$sub_genre)
unique(df$main_genre)
```


Checking for Outliers. 
Now let's have a look at the summary of our dataframe
```{r}
summary(df)
```
From summary, we can notice that max of acousticness is very high
let's try to understand this distribution better using box plot

```{r}
boxplot(df$acousticness)
```

Finding the rows where Acousticness is greater than 1
```{r}
subset(df, acousticness >= 1)%>%
  select(acousticness,everything())
```
As we know that acousticness can't be greater than 1, the values might be due to incorrect data we can divide the acousticness by 1000 and get values between 0 and 1

```{r}
df <- 
df %>% 
  mutate(acousticness = ifelse(acousticness>1,
                            acousticness/1000,
                            acousticness))
```

To verify the above operation.
```{r}
subset(df, acousticness >= 1)%>%
  select(acousticness,everything())
```

```{r}
summary(df)
```
Now Max of acousticness is below 1
```{r}
boxplot(df$acousticness)
```


From summary, we can also notice that max of liveness is very high
let's try to understand this distribution better using box plot

```{r}
boxplot(df$liveness)
```
We can see that Liveliness for few values are greater than 1. To fix these outliers, 
we are replacing it with mean.

Finding the rows where Acousticness is greater than 1
```{r}
subset(df, liveness >= 1)%>%
  select(liveness,everything())
```



```{r}
df <- 
df %>% 
    mutate(liveness = ifelse(liveness>1,
                             liveness/100,
                             liveness))
```
Plotting the box plot to verify the distribution.
```{r}
boxplot(df$liveness)
```

We see that there are still outliers. We are replacing it with the mean.
```{r}
df <- df %>% 
    mutate(liveness = ifelse(liveness>IQR(df$liveness) * 1.5,
                             mean(liveness), # liveness/100,
                             liveness))
```

Plotting the box plot to verify the distribution.
```{r}
boxplot(df$liveness)
```

For verifying the changes.
```{r}
subset(df, liveness >= 1)%>%
  select(liveness,everything())
```


Checking the summary to confirm the data for all columns.
```{r}
summary(df)
```


Saving the dataset to csv format.
```{r}
write.csv(df,"D:/MS_in_DS/Stats_1/Assignments/Assignment_1/Nikhil_BM_Stats_Assignment1/Stats_Assignment_1_Cleaned.csv",
         row.names = FALSE)
```
