---
title: "Spotify Project Clusterning"
author: "Prasenjeet Madhav Gadhe"
date: "2022-12-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(plotly)
library(corrplot)
library(factoextra)
library(plyr)
library(RColorBrewer)
library(funModeling)
library(knitr)
library(DataExplorer)
```

# Reading the clean Data
```{r}
df <- read.csv("C:/Users/jeetp/Desktop/Project/00.Spotify_Project/Stats_Assignment_1_Cleaned_2.csv")
head(df,2)
```
# Generating the Report
```{r}
DataExplorer::create_report(df)
```

# Converting genres into numeric value
```{r}
genr <- factor(df$main_genre)
df$genr_new <- unclass(genr)                         # Convert categories to numeric
df$genr_new  
#head(df,2)
```
## Model Building 
# Data Clustering
# We will be using specific attributes for clustering, scaling those attributes
```{r}
spotify_clu <- scale(df[,c(11,12,13,14,16,17,18,19,20,21,22)])
spotify_scaled <- cbind(spotify_clu, df$genr_new)

summary(spotify_scaled)
```

# Trying to find the optimal number of clusters using Elbow method

You can also embed plots, for example:

```{r}
wss <- function(data, maxCluster = 12) {
  SSw <- (nrow(data) - 1) * sum(apply(data, 2, var))
  SSw <- vector()
  for (i in 2:maxCluster) {
    SSw[i] <- sum(kmeans(data, centers = i)$withinss)
  }
  plot(1:maxCluster, SSw, type = "o", xlab = "Number of Clusters", ylab = "Within groups sum of squares", pch=19)
}

wss(spotify_scaled)
```

# In the plot as we can see at k =9 it appears to be the threshold point however, we can not se the abrupt point.


##K-mean Clustering
```{r}
spotify_kmeans <- kmeans(spotify_scaled, centers = 4)
spotify_kmeans$size
```

```{r}
spotify_kmeans$centers
```

```{r}
df$cluster <- spotify_kmeans$cluster
tail(df)
```
```{r}
fviz_cluster(spotify_kmeans, data=spotify_scaled)
```
```{r}
spotify_kmeans$tot.withinss
```

```{r}
spotify_kmeans$totss
```
```{r}
spotify_kmeans$betweenss
```
```{r}
((spotify_kmeans$betweenss)/(spotify_kmeans$totss))*100
```

```{r}
df %>% 
  group_by(cluster) %>% 
  summarise_all(mean) %>% 
  select(cluster, acousticness, danceability, energy, instrumentalness, speechiness, valence, liveness)
```


# Input Song Name and Artist Name
```{r}
df %>% 
  filter(song_name == "Stay", artist == "Hungry Lucy")
clust <- df$cluster[df$song_name == "Stay" & df$artist =="Hungry Lucy"]
clust
```

# Input for Recommended genre and Clusrter
```{r}
df %>% 
  filter(cluster == clust) %>% 
  sample_n(5)
```
